Towards more realistic model representation of microbial population dynamics in complex environment

### 2020-03-30 SI_code_revised.Rmd
Code revision - removing community data and Monod plus model, simplyfing Sub-microbial model

DNA and protein concentrations are in $\mu$grams per incubation vial. Both values are recalculated to unit of carbon per one ml of media in the vial (4 ml) using the conversion factors reported in Vrede et al. (2004).

@@

```
##Statistical analysis
Linear Mixed-Effect models are used to evaluate the effect of substrate and the type of physical barrier on measured respiration rate, DNA and protein concentration. Becuase all three variables display significant variability over time, the time of measurement is treated as a random effect. Type of the barrier is nested within the substrate.

@@

```
$~$
$~$
$~$
Next, the effect of the substrate and the presence of a physical barrier on microbial community structure and diversity is evaluated. First, the bray distance indices are calculated on a normalized data. Then, Permutational Multivariate Analysis of Variance is used to estimate the treatment effects. Analysis permutations are constrained to an individual times of incubation to avoid the potential confounding effect of time.

```{r vegan analysis, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#Bacteria
bac.norm<-decostand(bac2, method=c("normalize"))
bac.dist<-vegdist(bac.norm, method = "bray")
bac.ad<-adonis(bac.dist~Substrate+Structure/Substrate, bac_env, strata = bac_env$Day)

#Fungi
fungi.norm<-decostand(fungi2, method=c("normalize"))
fungi.dist<-vegdist(fungi.norm, method = "bray")
fungi.ad<-adonis(fungi.dist~Substrate+Structure/Substrate, fungi_env, strata = fungi_env$Day)

t.s01<-cbind(as.data.frame(bac.ad$aov[c(4, 1, 6)])[c(1:3), ], as.data.frame(fungi.ad$aov[c(4, 1, 6)])[c(1:3), ])

row.names(t.s01)<-c("Substrate", "Structure", "Substrate*Structure")

kable(t.s01, digits = c(2, 1, 3, 2, 1, 3), align = 'c', "html", booktabs=T, col.names = c("F", "df", "p", "F", "df", "p"), row.names=T, caption = "_**Table S2:** Results of permutational multivariate analysis of variance. Effect of substrate, the type of physical barrier (Structure) and their interaction was evaluated. Statistical analysis was conducted for bacterial and fungal part of the microbial community separately. F values, degrees of freedom (df) and the probability (p) are reported._") %>% 
  kable_styling() %>% 
  add_header_above(c(" " = 1, "Bacteria" = 3, "Fungi" = 3),
                   italic=T)

```
$~$
$~$
$~$
$~$
$~$
$~$
```{r community figure, echo=FALSE, fig.height=5, fig.width=10, fig.cap="Fig. S2: Multivariate distances between six experimental groups that combine two different substrate treatments (Glucose or Cellobiose) and three different types of physical barrier (BROTH, WOOL, GLASS - black, grey and empty symbols respectively) in respect to structure of bacterial ((a); 16S) and fungal ((b); ITS) part of microbial community."}
#Bacteria
bacm<-as.matrix(bac.dist)
#Intercept - Glucose - BROTH
mGB<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Glucose"))]
mGB[mGB==0]<-NA
#Glucose - GLASS
mGG<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="GLASS"),which(bac_env$Substrate=="Glucose"))]
mGG[mGG==0]<-NA
#Glucose - WOOL
mGW<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="WOOL"),which(bac_env$Substrate=="Glucose"))]
mGW[mGW==0]<-NA
#Cellobiose - BROTH
mCB<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose"))]
mCB[mCB==0]<-NA
#Cellobiose - GLASS
mCG<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="GLASS"),which(bac_env$Substrate=="Cellobiose"))]
mCG[mCG==0]<-NA
#Cellobiose - WOOL
mCW<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="WOOL"),which(bac_env$Substrate=="Cellobiose"))]
mCW[mCW==0]<-NA

bacdiff<-data.frame(Dist=c(mean(mGB, na.rm=T), mean(mGG, na.rm=T), mean(mGW, na.rm=T), 0, mean(mCG, na.rm=T), mean(mCW, na.rm=T)),
                    Dist.sd=c(sd(mGB, na.rm=T), sd(mGG, na.rm=T), sd(mGW, na.rm=T), sd(mCB, na.rm=T), sd(mCG, na.rm=T), sd(mCW, na.rm=T)),
                    Substrate=c(rep("Glucose", 3), rep("Cellobiose", 3)),
                    Structure=rep(c("BROTH", "GLASS", "WOOL"), 2))

#Fungi
fm<-as.matrix(fungi.dist)
#Intercept - Glucose - BROTH
fGB<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Glucose"))]
fGB[fGB==0]<-NA
#Glucose - GLASS
fGG<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="GLASS"),which(fungi_env$Substrate=="Glucose"))]
fGG[fGG==0]<-NA
#Glucose - WOOL
fGW<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="WOOL"),which(fungi_env$Substrate=="Glucose"))]
fGW[fGW==0]<-NA
#Cellobiose - BROTH
fCB<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose"))]
fCB[fCB==0]<-NA
#Cellobiose - GLASS
fCG<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="GLASS"),which(fungi_env$Substrate=="Cellobiose"))]
fCG[fCG==0]<-NA
#Cellobiose - WOOL
fCW<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="WOOL"),which(fungi_env$Substrate=="Cellobiose"))]
fCW[fCW==0]<-NA

fdiff<-data.frame(Dist=c(mean(fGB, na.rm=T), mean(fGG, na.rm=T), mean(fGW, na.rm=T), 0, mean(fCG, na.rm=T), mean(fCW, na.rm=T)),
                    Dist.sd=c(sd(fGB, na.rm=T), sd(fGG, na.rm=T), sd(fGW, na.rm=T), sd(fCB, na.rm=T), sd(fCG, na.rm=T), sd(fCW, na.rm=T)),
                    Substrate=c(rep("Glucose", 3), rep("Cellobiose", 3)),
                    Structure=rep(c("BROTH", "GLASS", "WOOL"), 2))

bacplot<-ggplot(bacdiff, aes(Substrate, Dist))+
  geom_point(position = position_dodge(width = 0.5), aes(fill=Structure), shape=21, cex=6)+
  scale_fill_manual(values = c("black", "grey", "white"))+
  scale_colour_manual(values = c("black", "black", "black"))+
  geom_errorbar(aes(ymin=Dist-Dist.sd, ymax=Dist+Dist.sd, colour=Structure),
                position=position_dodge(width = 0.5), width=0.1)+
  ylab("16S Distance")+
  ggtitle("a)")+
  theme_min+theme(axis.title.x = element_blank(),
                  legend.title = element_blank(),
                  legend.position = c(0.8, 0.2))
fplot<-ggplot(fdiff, aes(Substrate, Dist))+
  geom_point(position = position_dodge(width = 0.5), aes(fill=Structure), shape=21, cex=6, show.legend = F)+
  scale_fill_manual(values = c("black", "grey", "white"))+
  scale_colour_manual(values = c("black", "black", "black"))+
  geom_errorbar(aes(ymin=Dist-Dist.sd, ymax=Dist+Dist.sd, colour=Structure),
                position=position_dodge(width = 0.5), width=0.1, show.legend = F)+
  ylab("ITS Distance")+
  ggtitle("b)")+
  theme_min+theme(axis.title.x = element_blank(),
                  legend.title = element_blank())

grid.arrange(bacplot, fplot, nrow=1)

$~$
$~$
$~$

##Microbial-explicit models analysis
###All models are first calibrated against respiration rate
Calibration of each model is done in four steps. First, the model is calibrated against data across all treatments. Second, the model is calibrated for each level of incubation system complexity separately. Third, the model is calibrated for Glucose and Cellobiose treatments separately. Fourth, the model is calibrated for each experimental treatment separately. 
Goodness of correspondence between measurements and model simulations are calculated across all treatments. By doing so, the effect of experimental treatment is evaluated. If the model simulation fits the data significantly better when model parameters are estimated for different treatments separately, effect of this treatment is significant.

@@

####Monod model
```{r monod respiration calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/monod_r.R")
source("../Rscripts/monod_r_fix.R")
#across all treatments
monod_r1<-monod_r(data=d, FACT = 4)
#########################################################
no_cors<-detectCores()
cl<-makeCluster(no_cors)
registerDoParallel(cl)
##########################################################
#for different structures
monod_r2<-monod_r(data=d, FACT = 2)
monod_r2_fix<-monod_r_fix(data=d, FACT = 2)
#for different substrates
monod_r3<-monod_r(data=d, FACT = 1)
monod_r3_fix<-monod_r_fix(data=d, FACT = 1)
#for each separately  
monod_r4<-monod_r(data=d, FACT = 3)
monod_r4_fix<-monod_r_fix(data=d, FACT = 3)
###################################################
stopImplicitCluster()
###################################################
# monod_r1$goodness
# monod_r2$goodness
# monod_r3$goodness
# monod_r4$goodness

@@

####Monod plus model
```{r Monod plus respiration calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/monodplus_r.R")
source("../Rscripts/monodplus_r_fix.R")
#across all treatments
monodplus_r1<-monodplus_r(data=d, FACT = 4)
##################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
##################################################
#for different structures
monodplus_r2<-monodplus_r(data=d, FACT = 2)
monodplus_r2_fix<-monodplus_r_fix(data=d, FACT = 2)
#for different substrates
monodplus_r3<-monodplus_r(data=d, FACT = 1)
monodplus_r3_fix<-monodplus_r_fix(data=d, FACT = 1)
#for each separately 
monodplus_r4<-monodplus_r(data=d, FACT = 3)
monodplus_r4_fix<-monodplus_r_fix(data=d, FACT = 3)
##################################################
stopImplicitCluster()
##################################################
monodplus_r1$goodness
monodplus_r2$goodness
monodplus_r3$goodness
monodplus_r4$goodness

# mean(c(monodplus_r4[[1]]$pars[["Cmic_0"]], monodplus_r4[[2]]$pars[["Cmic_0"]],
#        monodplus_r4[[3]]$pars[["Cmic_0"]], monodplus_r4[[4]]$pars[["Cmic_0"]],
#        monodplus_r4[[5]]$pars[["Cmic_0"]], monodplus_r4[[6]]$pars[["Cmic_0"]]))

monodplus_r1$goodness
monodplus_r2_fix$goodness
monodplus_r3_fix$goodness
monodplus_r4_fix$goodness
```

```{r Monod plus respiration calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#results of the previous iteration (not run)
# monodplus_r_results_par<-as.data.frame(rbind(monodplus_r1[[1]]$pars[c(1:5)],
#                     monodplus_r2_fix[[1]]$pars, monodplus_r2_fix[[2]]$pars,
#                     monodplus_r2_fix[[3]]$pars,
#                     monodplus_r3_fix[[1]]$pars, monodplus_r3_fix[[2]]$pars,
#                     monodplus_r4_fix[[1]]$pars, monodplus_r4_fix[[2]]$pars,
#                     monodplus_r4_fix[[3]]$pars, monodplus_r4_fix[[4]]$pars,
#                     monodplus_r4_fix[[5]]$pars, monodplus_r4_fix[[6]]$pars))
# monodplus_r_results_good<-as.data.frame(rbind(monodplus_r1$goodness,
#                                               monodplus_r2_fix$goodness,
#                                               monodplus_r3_fix$goodness,
#                                               monodplus_r4_fix$goodness))
# write.csv(monodplus_r_results_par, file = c("../monodplus_r_results_par.csv"), row.names = F)
# write.csv(monodplus_r_results_good, file = c("../monodplus_r_results_good.csv"), row.names = F)

monodplus_r_results_par<-read.csv("monodplus_r_results_par.csv")
monodplus_r_results_good<-read.csv("monodplus_r_results_good.csv")

t.s2.1<-data.frame(Vmax=monodplus_r_results_par[,1],
                 Km=monodplus_r_results_par[,2],
                 CUE=monodplus_r_results_par[,3],
                 mr=monodplus_r_results_par[,4],
                 k=monodplus_r_results_par[,5])
row.names(t.s2.1)<-c("All", "Broth", "Glass", "Wool",
                    "Celluloze", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose")

@@

kable(t.s2.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('v', 'k', '$\\Y_{CUE}$', 'h', 'm'), row.names=T, caption = "_**Table S5:** Monod plus model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monodplus_r_results_good[,c(4, 1, 2, 3)], digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S6:** Goodness of fit of Monnod plus model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

two_pool_r1$goodness
two_pool_r2_fix$goodness
two_pool_r3_fix$goodness
two_pool_r4_fix$goodness
```

#results of the previous iteration (not run)
# two_pool_r_results_par<-rbind(two_pool_r1[[1]]$pars[c(1:6)],
#                     two_pool_r2_fix[[1]]$pars, two_pool_r2_fix[[2]]$pars,
#                     two_pool_r2_fix[[3]]$pars,
#                     two_pool_r3_fix[[1]]$pars, two_pool_r3_fix[[2]]$pars,
#                     two_pool_r4_fix[[1]]$pars, two_pool_r4_fix[[2]]$pars,
#                     two_pool_r4_fix[[3]]$pars, two_pool_r4_fix[[4]]$pars,
#                     two_pool_r4_fix[[5]]$pars, two_pool_r4_fix[[6]]$pars)
# two_pool_r_results_good<-rbind(two_pool_r1$goodness,
#                                two_pool_r2$goodness,
#                                two_pool_r3$goodness,
#                                two_pool_r4$goodness)
# write.csv(two_pool_r_results_par, file = c("../two_pool_r_results_par.csv"), row.names = F)
# write.csv(two_pool_r_results_good, file = c("../two_pool_r_results_good.csv"), row.names = F)

t.s3.1<-data.frame(Vmax=two_pool_r_results_par[,1],
                 Km=two_pool_r_results_par[,2],
                 mo=two_pool_r_results_par[,3],
                 f=two_pool_r_results_par[,4],
                 k=two_pool_r_results_par[,5],
                 Yu=two_pool_r_results_par[,6])
row.names(t.s3.1)<-c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose")

kable(t.s3.1, digits = c(2, 2, 3, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('v', 'k', 'h', 'f', 'm', 'Y'), row.names=T, caption = "_**Table S7:** Sub-microbial model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

###All models are calibrated against respiration rate, proteins and DNA concentration in the cells.
All models are calibrated against all data we measured. In addition to model parameters reported in tables above, conversion factors between proteins/DNA concentration and microbial biomass are estimated. For Monod and Monod plus models, initial concentration of microbial biomass is defined as an initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and microbial biomass. For Sub-microbial model, initial concentration of Structures is defined similarly - i.e. initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and Structures. However, initial concetration of Reserves cannot be defined because initial concentration of proteins is not know. Therefore, one more parameter is defined. This parameter is denoted as RS_0 in the respective function and define initial concentration of Reserves per unit of Structures. This parameter has to be fixed across treatments to reflect the reality.
In addition, Monod and Monod plus models are calibrated against respiration rate and proteins or DNA concentration separately. When calibrated againnst protein concentration, initial concentration of microbial biomass is estimated and fixed across treatments.

@@

####Monod plus model
```{r Monod plus model calibartion, eval=F, echo=T}
#Proteins
#Loading the function 
source("../Rscripts/monodplus_prot.R")
source("../Rscripts/monodplus_prot_fix.R")
#across all treatments
monodplus_prot1<-monodplus_prot(data=d, FACT = 4)
###########################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###########################################################
#for different structures
monodplus_prot2<-monodplus_prot(data=d, FACT = 2)
monodplus_prot2_fix<-monodplus_prot_fix(data=d, FACT = 2)
#for different substrates
monodplus_prot3<-monodplus_prot(data=d, FACT = 1)
monodplus_prot3_fix<-monodplus_prot_fix(data=d, FACT = 1)
#for each separately 
monodplus_prot4<-monodplus_prot(data=d, FACT = 3)
monodplus_prot4_fix<-monodplus_prot_fix(data=d, FACT = 3)
###########################################################
stopImplicitCluster()
###########################################################
monodplus_prot1$goodness
monodplus_prot2_fix$goodness
monodplus_prot3_fix$goodness
monodplus_prot4_fix$goodness
###########################################################
#DNA
#Loading the function 
source("../Rscripts/monodplus_DNA.R")
#across all treatments
monodplus_DNA1<-monodplus_DNA(data=d, FACT = 4)
###########################################################
no_cors<-detectCores()-2
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###########################################################
#for different structures
monodplus_DNA2<-monodplus_DNA(data=d, FACT = 2)
#for different substrates
monodplus_DNA3<-monodplus_DNA(data=d, FACT = 1)
#for each separately 
monodplus_DNA4<-monodplus_DNA(data=d, FACT = 3)
###########################################################
stopImplicitCluster()
###########################################################
monodplus_DNA1$goodness
monodplus_DNA2$goodness
monodplus_DNA3$goodness
monodplus_DNA4$goodness
###########################################################
#DNA and Proteins
#Loading the function 
source("../Rscripts/monodplus_all.R")
#across all treatments
monodplus_all1<-monodplus_all(data=d, FACT = 4)
###########################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###########################################################
#for different structures
monodplus_all2<-monodplus_all(data=d, FACT = 2)
#for different substrates
monodplus_all3<-monodplus_all(data=d, FACT = 1)
#for each separately 
monodplus_all4<-monodplus_all(data=d, FACT = 3)
###########################################################
stopImplicitCluster()
###########################################################
monodplus_all1$goodness
monodplus_all2$goodness
monodplus_all3$goodness
monodplus_all4$goodness
```

```{r Monod plus overall calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# #results of the previous iteration (not run)
# ##All measurements together
# monodplus_all_results_par<-rbind(monodplus_all1[[1]]$pars, 
#                             monodplus_all2[[1]]$pars, monodplus_all2[[2]]$pars,
#                             monodplus_all2[[3]]$pars, 
#                             monodplus_all3[[1]]$pars, monodplus_all3[[2]]$pars,
#                             monodplus_all4[[1]]$pars, monodplus_all4[[2]]$pars,
#                             monodplus_all4[[3]]$pars, monodplus_all4[[4]]$pars,
#                             monodplus_all4[[5]]$pars, monodplus_all4[[6]]$pars)
# 
# monodplus_all_results_good<-rbind(monodplus_all1$goodness[c(2, 3, 1), ],
#                              monodplus_all2$goodness[c(2, 3, 1), ],
#                              monodplus_all3$goodness[c(2, 3, 1), ],
#                              monodplus_all4$goodness[c(2, 3, 1), ])
# 
# write.csv(monodplus_all_results_par, file = c("../monodplus_all_results_par.csv"), row.names = F)
# write.csv(monodplus_all_results_good, file = c("../monodplus_all_results_good.csv"), row.names = F)
# 
# ##Respiration rate and proteins
# monodplus_prot_results_par<-rbind(monodplus_prot1[[1]]$pars[c(1:6)], 
#                             monodplus_prot2_fix[[1]]$pars, monodplus_prot2_fix[[2]]$pars,
#                             monodplus_prot2_fix[[3]]$pars, 
#                             monodplus_prot3_fix[[1]]$pars, monodplus_prot3_fix[[2]]$pars,
#                             monodplus_prot4_fix[[1]]$pars, monodplus_prot4_fix[[2]]$pars,
#                             monodplus_prot4_fix[[3]]$pars, monodplus_prot4_fix[[4]]$pars,
#                             monodplus_prot4_fix[[5]]$pars, monodplus_prot4_fix[[6]]$pars)
# 
# monodplus_prot_results_good<-rbind(monodplus_prot1$goodness[c(2,1), ],
#                              monodplus_prot2_fix$goodness[c(2,1), ],
#                              monodplus_prot3_fix$goodness[c(2,1), ],
#                              monodplus_prot4_fix$goodness[c(2,1), ])
# 
# write.csv(monodplus_prot_results_par, file = c("../monodplus_prot_results_par.csv"), row.names = F)
# write.csv(monodplus_prot_results_good, file = c("../monodplus_prot_results_good.csv"), row.names = F)
# 
# ##Respiration rate and DNA
# monodplus_DNA_results_par<-rbind(monodplus_DNA1[[1]]$pars, 
#                             monodplus_DNA2[[1]]$pars, monodplus_DNA2[[2]]$pars,
#                             monodplus_DNA2[[3]]$pars, 
#                             monodplus_DNA3[[1]]$pars, monodplus_DNA3[[2]]$pars,
#                             monodplus_DNA4[[1]]$pars, monodplus_DNA4[[2]]$pars,
#                             monodplus_DNA4[[3]]$pars, monodplus_DNA4[[4]]$pars,
#                             monodplus_DNA4[[5]]$pars, monodplus_DNA4[[6]]$pars)
#  
# monodplus_DNA_results_good<-rbind(monodplus_DNA1$goodness[c(2, 1), ],
#                              monodplus_DNA2$goodness[c(2, 1), ],
#                              monodplus_DNA3$goodness[c(2, 1), ],
#                              monodplus_DNA4$goodness[c(2, 1), ])
# 
# write.csv(monodplus_DNA_results_par, file = c("../monodplus_DNA_results_par.csv"), row.names = F)
# write.csv(monodplus_DNA_results_good, file = c("../monodplus_DNA_results_good.csv"), row.names = F)

#All measurements
monodplus_all_results_par<-read.csv("monodplus_all_results_par.csv")
monodplus_all_results_good<-read.csv("monodplus_all_results_good.csv")
monodplus_all_results_good$variable<-rep(c("Respiration", "Proteins", "DNA"), 4)
monodplus_all_results_good[monodplus_all_results_good$R2<0, "R2"]<-0
#Respiration and proteins
monodplus_prot_results_par<-read.csv("monodplus_prot_results_par.csv")
monodplus_prot_results_good<-read.csv("monodplus_prot_results_good.csv")
monodplus_prot_results_good$variable<-rep(c("Respiration", "Proteins"), 4)
monodplus_prot_results_good[monodplus_prot_results_good$R2<0, "R2"]<-0
#Respiration and DNA
monodplus_DNA_results_par<-read.csv("monodplus_DNA_results_par.csv")
monodplus_DNA_results_good<-read.csv("monodplus_DNA_results_good.csv")
monodplus_DNA_results_good$variable<-rep(c("Respiration", "DNA"), 4)
monodplus_DNA_results_good[monodplus_DNA_results_good$R2<0, "R2"]<-0

>>>>

DNA and protein concentrations are in $\mu$grams per incubation vial. Both values are converted to unit of carbon per one ml of media in the vial (4 ml) using the conversion factors reported in Vrede et al. (2004).

@@

```
##Statistical analysis
Linear Mixed-Effect models are used to evaluate the effect of substrate and the type of physical barrier on measured respiration rate, DNA and protein concentration. Because all three variables display significant variability over time, the time of measurement is treated as a random effect factor. Type of the barrier is nested within the substrate.

@@

```
$~$
$~$
$~$
##Microbial-explicit models analysis
###Both models are first calibrated against respiration rate
Calibration of both models is done in four steps. First, the model is calibrated against data across all treatments. Second, the model is calibrated for each level of incubation system complexity separately. Third, the model is calibrated for Glucose and Cellobiose treatments separately. Fourth, the model is calibrated for each experimental treatment separately. 
Goodness of correspondence between measurements and model simulations are calculated across all treatments. By doing so, the effect of experimental treatment is evaluated. If the model simulation fits the data significantly better when model parameters are estimated for different treatments separately, effect of this treatment is significant.

@@

####Monod model
```{r monod respiration calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/monod_r.R")
source("../Rscripts/monod_r_fix.R")
#across all treatments
monod_r1<-monod_r(data=d, FACT = 4)
#########################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
##########################################################
#for different structures
monod_r2<-monod_r(data=d, FACT = 2)
monod_r2_fix<-monod_r_fix(data=d, FACT = 2)
#for different substrates
monod_r3<-monod_r(data=d, FACT = 1)
monod_r3_fix<-monod_r_fix(data=d, FACT = 1)
#for each separately  
monod_r4<-monod_r(data=d, FACT = 3)
monod_r4_fix<-monod_r_fix(data=d, FACT = 3)
###################################################
stopImplicitCluster()
###################################################
# monod_r1$goodness
# monod_r2$goodness
# monod_r3$goodness
# monod_r4$goodness

@@

two_pool_r1$goodness
two_pool_r2_fix$goodness
two_pool_r3_fix$goodness
two_pool_r4_fix$goodness

#results of the previous iteration (not run)
# two_pool_r_results_par<-rbind(two_pool_r1[[1]]$pars[c(1:5)],
#                    two_pool_r2_fix[[1]]$pars, two_pool_r2_fix[[2]]$pars,
#                    two_pool_r2_fix[[3]]$pars,
#                    two_pool_r3_fix[[1]]$pars, two_pool_r3_fix[[2]]$pars,
#                    two_pool_r4_fix[[1]]$pars, two_pool_r4_fix[[2]]$pars,
#                    two_pool_r4_fix[[3]]$pars, two_pool_r4_fix[[4]]$pars,
#                    two_pool_r4_fix[[5]]$pars, two_pool_r4_fix[[6]]$pars)
# two_pool_r_results_good<-rbind(two_pool_r1$goodness,
#                               two_pool_r2$goodness,
#                               two_pool_r3$goodness,
#                               two_pool_r4$goodness)
# write.csv(two_pool_r_results_par, file = c("../two_pool_r_results_par.csv"), row.names = F)
# write.csv(two_pool_r_results_good, file = c("../two_pool_r_results_good.csv"), row.names = F)

t.s3.1<-data.frame(Vmax=two_pool_r_results_par[,1],
                 Km=two_pool_r_results_par[,2],
                 f=two_pool_r_results_par[,3],
                 k=two_pool_r_results_par[,4],
                 Yu=two_pool_r_results_par[,5])
row.names(t.s3.1)<-c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose")

@@

kable(t.s3.1, digits = c(2, 2, 2, 3, 2), align = 'c', "html", booktabs=T, col.names = c('v', 'k', 'f', 'm', 'Y'), row.names=T, caption = "_**Table S7:** Sub-microbial model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

@@

###Both models are calibrated against respiration rate, proteins and DNA concentration in the cells.
Both models are calibrated against all data we measured. In addition to model parameters reported in tables above, conversion factors between proteins/DNA concentration and microbial biomass are estimated. For Monod model, initial concentration of microbial biomass is defined as an initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and microbial biomass. For Sub-microbial model, initial concentration of Structures is defined similarly - i.e. initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and Structures. However, initial concetration of Reserves cannot be defined because initial concentration of proteins is not know. Therefore, one more parameter is defined. This parameter is denoted as RS_0 in the respective function and define initial concentration of Reserves per unit of Structures. This parameter has to be fixed across treatments to reflect the reality.
In addition, Monod model is calibrated against respiration rate and proteins or DNA concentration separately. When calibrated againnst protein concentration, initial concentration of microbial biomass is estimated and fixed across treatments.

@@

####Sub-microbial model
Not all model parameters vary significantly across treatments. Parameters that doesn't show any variability across treatments are fixed.
```{r Sub-microbial model calibration, eval=F, echo=T}
#Loading the function 
source("./Rscripts/two_pool_allx.R")
source("../Rscripts/two_pool_allx_fix.R")
#across all treatments
two_pool_allx_out1<-two_pool_allx(data=d, FACT = 4)

###########################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###########################################################
#for different structures
two_pool_allx_out2<-two_pool_allx(data=d, FACT = 2)

two_pool_allx_fix_out2<-two_pool_allx_fix(data=d, FACT = 2)
two_pool_allx_fix_out2$goodness

two_pool_allx_fix_out2[[1]]$pars
two_pool_allx_fix_out2[[2]]$pars
two_pool_allx_fix_out2[[3]]$pars

# mean(c(two_pool_allx_fix_out2[[1]]$pars[["Vmax"]],
#        two_pool_allx_fix_out2[[2]]$pars[["Vmax"]],
#        two_pool_allx_fix_out2[[3]]$pars[["Vmax"]]))
# mean(c(two_pool_allx_fix_out2[[1]]$pars[["Yu"]],
#        two_pool_allx_fix_out2[[2]]$pars[["Yu"]],
#        two_pool_allx_fix_out2[[3]]$pars[["Yu"]]))
# mean(c(two_pool_allx_fix_out2[[1]]$pars[["k"]],
#        two_pool_allx_fix_out2[[2]]$pars[["k"]],
#        two_pool_allx_fix_out2[[3]]$pars[["k"]]))
############################################################################
#in the following code, the Vmax, Ys and k parameters are fixed across treatments
#becuase its variability across treatments is minimal (data not shown)
source("./Rscripts/two_pool_all_fixVmaxYuk.R")
two_pool_all_fixVmaxYuk_out2<-two_pool_all_fixVmaxYuk(data=d, FACT = 2)
two_pool_all_fixVmaxYuk_out2$goodness

############################################################################
#for different substrates
two_pool_allx_fix_out3<-two_pool_allx_fix(data=d, FACT = 1)
two_pool_allx_fix_out3$goodness
#for each separately 
two_pool_allx_fix_out4<-two_pool_allx_fix(data=d, FACT = 3)
two_pool_allx_fix_out4$goodness
#in the following code, the Vmax Ys, and k parameters are fixed across treatments
two_pool_all_fixVmaxYuk_out4<-two_pool_all_fixVmaxYuk(data=d, FACT = 3)
two_pool_all_fixVmaxYuk_out4$goodness
###########################################################
stopImplicitCluster()
###########################################################
two_pool_allx_out1$goodness
two_pool_all_fixVmaxYuk_out2$goodness
two_pool_allx_fix_out3$goodness
two_pool_all_fixVmaxYuk_out4$goodness

# mean(c(two_pool_allx_out4[[1]]$pars[["RS_0"]],
#        two_pool_allx_out4[[2]]$pars[["RS_0"]],
#        two_pool_allx_out4[[3]]$pars[["RS_0"]],
#        #two_pool_allx_out4[[4]]$pars[["RS_0"]],
#        two_pool_allx_out4[[5]]$pars[["RS_0"]],
#        two_pool_allx_out4[[6]]$pars[["RS_0"]]))
